\Large{Adversarial Variational Optimization of Non-Differentiable Simulators}

### 2017-07-14 tex/paper.tex
save

Algorithm~\ref{alg:avo} represents the simplest version of AVO; however, the
variance of the noisy estimator of the gradients may be too large to be useful
in many problems. Variance reduction techniques such as Rao-Blackwellization,
Control Variates, and quasi-Monte Carlo can be applied here~\citep{ranganath2014black, tran2017variational}.
Similarly, various fast
algorithms for calculating the exact (entropically regularized) Wasserstein distance on
empirical distributions are being developed~\citep{cuturi2013sinkhorn, genevay2016stochastic, montavon2016wasserstein}.

@@

Typically in variational inference one starts with a tractable prior
$p(\bftheta)$ and a tractable likelihood $p(\bfx |\bftheta )$, thus fixing the
posterior $p(\bftheta | \bfx )$.   After optimizing $q(\bftheta | \bfpsi)$ to
approximate the posterior, one would criticize the model through the the
posterior predictive distribution as advocated by Box~\citep{box1980sampling}.
Effectively, we are optimizing the variational posterior $q(\bftheta | \bfpsi)$ by
directly minimizing the  divergence of the corresponding posterior predictive
distribution $\qxpsi$ and the data distribution $p_r(\bfx)$ without evaluating
the intractable likelihood $p(\bfx | \bftheta )$ or specifying a prior.
Consequently, the proposed algorithm is a form of variational inference through
model criticism~\citep{box1980sampling}, which we denote AVO-EB. While the
likelihood is intractable, it is fixed, thus one can view the optimization of
$\bfpsi$ through the lens of empirical Bayes, though the prior family
$p(\bftheta | \bfpsi)$ is not explicitly specified.

@@

\kcnote{Clarify hierarchical models and iid data x. Factor of $N/M$ for AVO-EB?}

>>>>

Algorithm~\ref{alg:avo} represents the simplest version of AVO; however, the
variance of the noisy estimator of the gradients may be too large to be useful in many problems. Improvements are possible using variance reduction techniques such as Rao-Blackwellization,
Control Variates, and quasi-Monte Carlo~\citep{ranganath2014black, tran2017variational}. 
Similarly, various fast algorithms for calculating the exact (entropically regularized) Wasserstein distance on
empirical distributions are being developed~\citep{cuturi2013sinkhorn, genevay2016stochastic, montavon2016wasserstein}. 
Similarly, the algorithm could be adopted to use estimates to the KL distance~\citep{perez2009estimation} without the adversarial optimization for the critic. These variants may be particularly useful for small $N$.

@@

Typically in variational inference one starts with a tractable prior
$p(\bftheta)$ and a tractable likelihood $p(\bfx |\bftheta )$, thus fixing the
posterior $p(\bftheta | \bfx )$.   After optimizing $q(\bftheta | \bfpsi)$ to
approximate the posterior, one would criticize the model through the the
posterior predictive distribution as advocated by Box~\citep{box1980sampling}.
Effectively, we are optimizing the variational posterior $q(\bftheta | \bfpsi)$ by
directly minimizing the  divergence of the corresponding posterior predictive
distribution $\qxpsi$ and the data distribution $p_r(\bfx)$ without evaluating
the intractable likelihood $p(\bfx | \bftheta )$ or specifying a prior.
Consequently, the proposed algorithm is a form of variational inference through
model criticism~\citep{box1980sampling}.

We can also view the optimization  through the lens of empirical Bayes (EB),
where one specifies a family of priors and $p(\bftheta | \bfpsi)$ and then estimates $\bfpsi^*$ using the data.
%In our case, we find the $\bfpsi^*$ that minimizes the Wasserstein distance between the data distribution and the 
%posterior predictive distribution. 
Traditionally, we would have access to the likelihood and optimize with respect to the family of posteriors
 $p(\bftheta | \bfx, \bfpsi) = p(\bfx | \bftheta) p(\bftheta | \bfpsi^*) / \int p(\bfx | \bftheta) p(\bftheta | \bfpsi^*) d\bftheta $
as an intermediate step to computing $\bfpsi^*$. In our case, the likelihood is intractable and we have not explicitly 
specified a prior family. However, the procedure corresponds to empirical Bayes if we identify 
$q(\bftheta | \bfpsi^*) = p(\bftheta | \bfx , \bfpsi)$. Thus we call our method AVO-EB and note that it is applicable even outside of the likelihood-free context.

Algorithm~\ref{alg:avo}, can be easily modified to have different mini-batch $M_r$ for the observed data and $M_g$ for the simulator. 
When $M_g \gg M_r$, then the variance of $\bftheta^*$ goes like $M_r$~\citep{bernton2017inference}. A final $\sqrt{N/M}$ scaling to the log-posterior $q(\bftheta | \bfpsi)$ is needed to reflect the full strength of the data.

%While the
%likelihood is intractable, it is fixed, thus one can view the optimization of
%$\bfpsi$ through the lens of empirical Bayes, though the prior family
%$p(\bftheta | \bfpsi)$ is not explicitly specified, 
%ie. $q(\bftheta | \bfpsi^*) = p(\bfx | \bftheta) p(\bftheta | \bfpsi^*) / \int d\bftheta p(\bfx | \bftheta) p(\theta | \bfpsi^*)$.

@@
